---
title: "stastics_pipeline"
author: "Zhang Linfeng"
date: "2022/9/9"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---


##library
```{r}
#library(tidyverse)
library(vegan)
library(reshape2)
library(dplyr)
library(ggplot2)
library(stringr)
library(phyloseq)
library(RColorBrewer)
library(tidyr)
library(ggsignif)
library(ggpubr)
```



##PERMANOVA

###univariate PERMANOVA
```{r}
metadata = read.csv('F:/ZLF/CAP/data/relative_data/metadata-all-220825.csv',row.names = 1)
df = read.csv("F:/ZLF/CAP/data/relative_data/dfall-220617.csv",row.names=1)
vals = c('onsetdays_new','sex','age','BMI', 'city',
         'smoke','basic_LRT_disease','pre_biotic','pre_immunosuppressive',
         'respiratory.support.invasive','death_case1','severe_case') 
tot.df = data.frame(R2 = rep(NA,length(vals)),
                    pval = rep(NA,length(vals)))
row.names(tot.df) <- vals
sub.meta = filter(metadata,d == 1)
for (val in vals){
  d1.meta = sub.meta[!is.na(sub.meta[,val]),]
  rownames(d1.meta) <- d1.meta$raw_id
  d1.df = df[,rownames(d1.meta)] %>% t() %>% as.data.frame()
  #View(sub.df)
  OTU = phyloseq::otu_table(d1.df, taxa_are_rows = F)
  physeq = phyloseq(OTU)
  jsd=phyloseq::distance(physeq, method = "jsd")
  
  res = adonis(as.formula(paste0("jsd~",val)),
               data = d1.meta, permutations = 999)
  
  pval = res$aov.tab$`Pr(>F)`[1]
  r2 = res$aov.tab$R2[1]
  tot.df[val,'pval'] = pval
  tot.df[val,'R2'] = r2
}
single_res <- tot.df
single_res$padj <- p.adjust (single_res$pval, method = 'fdr', n = (length(vals)))
single_res$group <- 'single'
single_res <- single_res[order(single_res[,'R2'],decreasing = T),]
single_res$value <- rownames(single_res)
single_res
```

### multivariate PERMANOVA
```{r}
metadata = read.csv('F:/ZLF/CAP/data/relative_data/metadata-all-220825.csv',row.names = 1)
df = read.csv("F:/ZLF/CAP/data/relative_data/dfall-220617.csv",row.names=1)
vals = c('age', 'city','respiratory.support.invasive','death_case1','severe_case') 
vals1 = c('onsetdays_new','sex','age','BMI', 'city',
         'smoke','basic_LRT_disease','pre_biotic','pre_immunosuppressive',
         'respiratory.support.invasive','death_case1','severe_case') 
tot.df = data.frame(R2 = rep(NA,length(vals)),
                    pval = rep(NA,length(vals)))
row.names(tot.df) <- vals
sub.meta = filter(metadata,d == 1)
for (val in vals) {sub.meta <- sub.meta[!is.na(sub.meta[,val]),]}
rownames(sub.meta) <- sub.meta$raw_id
sub.df = df[,rownames(sub.meta)] %>% t() %>% as.data.frame()
nrow(sub.meta)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res_11 <- adonis2(as.formula(paste0("jsd~",'age+city+respiratory.support.invasive+death_case1+severe_case')),
                  data = sub.meta,permutations = 999,by = 'margin')
res_11
tot.df$R2 = res_11$R2[1:5]
tot.df$pval = res_11$`Pr(>F)`[1:5]
multi_res <- tot.df
multi_res$padj <- p.adjust (multi_res$pval, method = 'fdr', n = (length(vals)))
multi_res$group <- 'multi'
multi_res <- multi_res[order(multi_res[,'R2'],decreasing = T),]
multi_res$value = rownames(multi_res)
a <- single_res[setdiff(rownames(single_res),rownames(multi_res)),]
a$R2 = 0
a$pval = 0
a$padj = 0
a$group = 'multi'
multi_res <- rbind(multi_res,a)
multi_res
```



###pathogen PERMANOVA

###pathogen data import
```{r}
#pathogens = c('Mycoplasma.pneumoniae','Klebsiella.pneumoniae','Influenza.A')
sub.meta <- filter(metadata, dfirst == 'dfirst') %>% filter(type %in% c('Mycoplasma.pneumoniae','Klebsiella.pneumoniae','Influenza.A'))
sub.df <- df[,rownames(sub.meta)] %>% t() %>% as.data.frame()
demy_df<-sub.df[rownames(filter(sub.meta,type == 'Mycoplasma.pneumoniae')),]
dekle_df<-sub.df[rownames(filter(sub.meta,type == 'Klebsiella.pneumoniae')),]
demy_df[,'g__Mycoplasma'] = 0
dekle_df[,'f__Enterobacteriaceae'] = 0
other_df <- sub.df[rownames(filter(sub.meta,type != 'Mycoplasma.pneumoniae' & type != 'Klebsiella.pneumoniae')),]
df_deself <- rbind(demy_df,dekle_df) %>% rbind(other_df) %>% decostand('total',MARGIN = 1)
metahea <- filter(metadata,severe_case == 'healthy')
dfhea <- df[,rownames(metahea)] %>% t() %>% as.data.frame()
df_deself_hea <- rbind(dfhea,df_deself)
meta_deself_hea <- metadata[rownames(df_deself_hea),] %>% select(raw_id,city,severe_case,Shannon,Mycoplasma.pneumoniae,Klebsiella.pneumoniae,Influenza.A)
meta_deself_hea <- left_join(meta_deself_hea,select(sub.meta,raw_id,type))
meta_deself_hea[is.na(meta_deself_hea[,'type']),'type'] <- 'healthy'
rownames(meta_deself_hea) <- meta_deself_hea$raw_id
# Bacteria Virus Mix
sub.meta <- filter(metadata, dfirst == 'dfirst')
sub.meta <- sub.meta[sub.meta[,'pathogen_type'] != 'na',]
sub.df <- df[,rownames(sub.meta)] %>% t() %>% as.data.frame()
metahea <- filter(metadata,severe_case == 'healthy')
dfhea <- df[,rownames(metahea)] %>% t() %>% as.data.frame()
df_pathogen_hea <- rbind(dfhea,sub.df)
meta_pathogen_hea <- metadata[rownames(df_pathogen_hea),] %>% select(raw_id,city,severe_case,Shannon)
meta_pathogen_hea <- left_join(meta_pathogen_hea,select(sub.meta,raw_id,pathogen_type))
meta_pathogen_hea[is.na(meta_pathogen_hea[,'pathogen_type']),'pathogen_type'] <- 'healthy'
rownames(meta_pathogen_hea) <- meta_pathogen_hea$raw_id
```


###pathogen PERMANOVA

###pathogen data import
```{r}
pathogens = c('Mycoplasma.pneumoniae','Klebsiella.pneumoniae','Influenza.A')
sub.meta <- filter(metadata, dfirst == 'dfirst') %>% filter(type %in% pathogens)
sub.df <- df[,rownames(sub.meta)] %>% t() %>% as.data.frame()
demy_df<-sub.df[rownames(filter(sub.meta,type == 'Mycoplasma.pneumoniae')),]
dekle_df<-sub.df[rownames(filter(sub.meta,type == 'Klebsiella.pneumoniae')),]
demy_df[,'g__Mycoplasma'] = 0
dekle_df[,'f__Enterobacteriaceae'] = 0
other_df <- sub.df[rownames(filter(sub.meta,type != 'Mycoplasma.pneumoniae' & type != 'Klebsiella.pneumoniae')),]
df_deself <- rbind(demy_df,dekle_df) %>% rbind(other_df) %>% decostand('total',MARGIN = 1)
metahea <- filter(metadata,severe_case == 'healthy')
dfhea <- df[,rownames(metahea)] %>% t() %>% as.data.frame()
df_deself_hea <- rbind(dfhea,df_deself)
meta_deself_hea <- metadata[rownames(df_deself_hea),] %>% select(raw_id,city,severe_case,Shannon,Mycoplasma.pneumoniae,Klebsiella.pneumoniae,Influenza.A)
meta_deself_hea <- left_join(meta_deself_hea,select(sub.meta,raw_id,type))
meta_deself_hea[is.na(meta_deself_hea[,'type']),'type'] <- 'healthy'
rownames(meta_deself_hea) <- meta_deself_hea$raw_id
# Bacteria Virus Mix
sub.meta <- filter(metadata, dfirst == 'dfirst')
sub.meta <- sub.meta[sub.meta[,'pathogen_type'] != 'na',]
sub.df <- df[,rownames(sub.meta)] %>% t() %>% as.data.frame()
metahea <- filter(metadata,severe_case == 'healthy')
dfhea <- df[,rownames(metahea)] %>% t() %>% as.data.frame()
df_pathogen_hea <- rbind(dfhea,sub.df)
meta_pathogen_hea <- metadata[rownames(df_pathogen_hea),] %>% select(raw_id,city,severe_case,Shannon)
meta_pathogen_hea <- left_join(meta_pathogen_hea,select(sub.meta,raw_id,pathogen_type))
meta_pathogen_hea[is.na(meta_pathogen_hea[,'pathogen_type']),'pathogen_type'] <- 'healthy'
rownames(meta_pathogen_hea) <- meta_pathogen_hea$raw_id
```


```{r}

##PH
d1.meta<-meta_deself_hea
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

d1.meta[d1.meta[,'type']!='healthy','type']='pathogen'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)
print(paste0('PH R2 = ',res$aov.tab$R2[1],' PH p-value = ',res$aov.tab$`Pr(>F)`[1]))


##MKIH
d1.meta<-meta_deself_hea
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
              data = sub.meta,permutations = 999)

print(paste0('MKIH R2 = ',res$aov.tab$R2[1],' MKIH p-value = ',res$aov.tab$`Pr(>F)`[1]))



##MKI
d1.meta<-meta_deself_hea %>% filter(type != 'healthy')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('MKI R2 = ',res$aov.tab$R2[1],' MKI p-value = ',res$aov.tab$`Pr(>F)`[1]))


##MH
d1.meta<-meta_deself_hea %>% filter(type == 'healthy' | type == 'Mycoplasma.pneumoniae')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('MH R2 = ',res$aov.tab$R2[1],' MH p-value = ',res$aov.tab$`Pr(>F)`[1]))



##KH
d1.meta<-meta_deself_hea %>% filter(type == 'healthy' | type == 'Klebsiella.pneumoniae')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('KH R2 = ',res$aov.tab$R2[1],' KH p-value = ',res$aov.tab$`Pr(>F)`[1]))


##IH
d1.meta<-meta_deself_hea %>% filter(type == 'healthy' | type == 'Influenza.A')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('IH R2 = ',res$aov.tab$R2[1],' IH p-value = ',res$aov.tab$`Pr(>F)`[1]))



##MK
#pathogens
#"Mycoplasma.pneumoniae" "Klebsiella.pneumoniae" "Influenza.A"  
d1.meta<-meta_deself_hea %>% filter(type == 'Mycoplasma.pneumoniae' | type == 'Klebsiella.pneumoniae')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('MK R2 = ',res$aov.tab$R2[1],' MK p-value = ',res$aov.tab$`Pr(>F)`[1]))


##MI
#pathogens
#"Mycoplasma.pneumoniae" "Klebsiella.pneumoniae" "Influenza.A"  
d1.meta<-meta_deself_hea %>% filter(type == 'Mycoplasma.pneumoniae' | type == 'Influenza.A')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('MI R2 = ',res$aov.tab$R2[1],' MI p-value = ',res$aov.tab$`Pr(>F)`[1]))

##KI
#pathogens
#"Mycoplasma.pneumoniae" "Klebsiella.pneumoniae" "Influenza.A"  
d1.meta<-meta_deself_hea %>% filter(type == 'Klebsiella.pneumoniae' | type == 'Influenza.A')
d1.df<-df_deself_hea
#d1.meta$respiratory.support_bin
val = 'type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'type')),
             data = sub.meta,permutations = 999)

print(paste0('KI R2 = ',res$aov.tab$R2[1],' KI p-value = ',res$aov.tab$`Pr(>F)`[1]))


##BVMH
d1.meta<-meta_pathogen_hea
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('BVMH R2 = ',res$aov.tab$R2[1],' BVMH p-value = ',res$aov.tab$`Pr(>F)`[1]))

##BVM
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type != 'healthy')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('BVM R2 = ',res$aov.tab$R2[1],' BVM p-value = ',res$aov.tab$`Pr(>F)`[1]))


##BH
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type == 'healthy' | pathogen_type == 'bacteria')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('BH R2 = ',res$aov.tab$R2[1],' BH p-value = ',res$aov.tab$`Pr(>F)`[1]))


##VH
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type == 'healthy' | pathogen_type == 'virus')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('VH R2 = ',res$aov.tab$R2[1],' VH p-value = ',res$aov.tab$`Pr(>F)`[1]))



##MH
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type == 'healthy' | pathogen_type == 'mix')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('MH R2 = ',res$aov.tab$R2[1],' MH p-value = ',res$aov.tab$`Pr(>F)`[1]))



##BV
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type == 'bacteria' | pathogen_type == 'virus')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('BV R2 = ',res$aov.tab$R2[1],' BV p-value = ',res$aov.tab$`Pr(>F)`[1]))


##BM
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type == 'bacteria' | pathogen_type == 'mix')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('BM R2 = ',res$aov.tab$R2[1],' BM p-value = ',res$aov.tab$`Pr(>F)`[1]))



##VM
d1.meta<-meta_pathogen_hea %>% filter(pathogen_type == 'virus' | pathogen_type == 'mix')
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]
sub.df = d1.df[rownames(sub.meta),]
#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('VM R2 = ',res$aov.tab$R2[1],' VM p-value = ',res$aov.tab$`Pr(>F)`[1]))



##PH

d1.meta<-meta_pathogen_hea 
d1.df<-df_pathogen_hea
#d1.meta$respiratory.support_bin
val = 'pathogen_type'

sub.meta = d1.meta[!is.na(d1.meta[,val]),]

sub.meta[sub.meta[,'pathogen_type'] != 'healthy',] = 'pathogen'

sub.df = d1.df[rownames(sub.meta),]

#View(sub.df)
OTU = phyloseq::otu_table(sub.df, taxa_are_rows = F)
physeq = phyloseq(OTU)
jsd=phyloseq::distance(physeq, method = "jsd")
res = adonis(as.formula(paste0("jsd~",'pathogen_type')),
             data = sub.meta,permutations = 999)

print(paste0('PH R2 = ',res$aov.tab$R2[1],' PH p-value = ',res$aov.tab$`Pr(>F)`[1]))

```



##cluster
```{r}
## PAM
pam1 <- function(x,k) list(cluster = pam(x,k, cluster.only=TRUE))

max_num = 20

nums=c(2:{max_num})
cord=df %>% t() %>% as.data.frame()

for(num in nums){{
  pam = pam1(cord,k=num)
  ofile=c("cord_pam_",num,".csv")
  write.csv(pam ,ofile)    
}}
```

##test in paper
###mechanical ventilation related CS in severity
```{r}
library(openxlsx)
transmeta = read.xlsx('F:/ZLF/CAP/data/relative_data/trans_meta1.xlsx',1,rowNames = T)
transmeta<-filter(transmeta,severity == 'yes')
transmeta_mv<-filter(transmeta,respiratory_support == '1' | respiratory_support == '2,1')
transmeta_oth<-filter(transmeta,respiratory_support != '1' & respiratory_support != '2,1')


trans<-data.frame(matrix(nrow = 2,ncol = 2))
trans<-data.frame(matrix(nrow = 2,ncol = 2))
rownames(trans)<-c('mv','others')
colnames(trans)<-c('self','to others')

trans[1,1]<-nrow(filter(transmeta_mv, d1 ==d5))
trans[1,2]<-nrow(filter(transmeta_mv, d1 !=d5))
trans[2,1]<-nrow(filter(transmeta_oth, d1 ==d5))
trans[2,2]<-nrow(filter(transmeta_oth, d1 !=d5))
fisher.test(trans)

for (i in 1:10) {
  trans<-data.frame(matrix(nrow = 2,ncol = 2))
  trans<-data.frame(matrix(nrow = 2,ncol = 2))
  colnames(trans)<-c('cs','csothers')
  rownames(trans)<-c('mv','non-mv')
  
  trans[1,1]<-nrow(filter(transmeta_mv, d1 != i & d5 == i))
  trans[1,2]<-nrow(filter(transmeta_mv, d1 !=i & d5 != i & d5 != d1))
  trans[2,1]<-nrow(filter(transmeta_oth, d1 != i & d5 == i))
  trans[2,2]<-nrow(filter(transmeta_oth, d1 !=i & d5 != i & d5 != d1))
  print(paste('cluter',i,'pvalue = ',fisher.test(trans)$p))
}

```

###intubated patients were transmitted to CS5 in Day 5 
```{r}
metadata = read.csv('F:/ZLF/CAP/data/relative_data/metadata-all-220617.csv',row.names = 1)
trans<-data.frame(matrix(nrow = 2,ncol = 2))
trans<-data.frame(matrix(nrow = 2,ncol = 2))
rownames(trans)<-c('intubate','unintubate')
colnames(trans)<-c('tocs5','tocsothers')

transmeta_mv<-filter(transmeta,respiratory_support == '1' | respiratory_support == '2,1')
transmeta_oth<-filter(transmeta,respiratory_support != '1' & respiratory_support != '2,1')
trans[1,1]<-filter(transmeta,d5 == 5) %>% 
  filter(respiratory_support == '1' | respiratory_support == '2,1') %>% nrow()
trans[2,1]<-filter(transmeta,d5 == 5) %>% 
  filter(respiratory_support != '1' & respiratory_support != '2,1') %>% 
   nrow()
trans[1,2]<-filter(transmeta,d5 != 5) %>% 
  filter(respiratory_support == '1' | respiratory_support == '2,1') %>% nrow()
trans[2,2]<-filter(transmeta,d5 != 5) %>% 
  filter(respiratory_support != '1' & respiratory_support != '2,1') %>% 
  nrow()
fisher.test(trans)
```

###intubated patients were transmitted to CS5 in Day 5 
```{r}

metadata = read.csv('F:/ZLF/CAP/data/relative_data/metadata-all-220617.csv',row.names = 1)
trans<-data.frame(matrix(nrow = 2,ncol = 2))
trans<-data.frame(matrix(nrow = 2,ncol = 2))
rownames(trans)<-c('intubate','unintubate')
colnames(trans)<-c('tocs5','tocsothers')

transmeta_mv<-filter(transmeta,respiratory_support == '1' | respiratory_support == '2,1')
transmeta_oth<-filter(transmeta,respiratory_support != '1' & respiratory_support != '2,1')
trans[1,1]<-filter(transmeta,d5 == 5) %>% 
  filter(respiratory_support == '1' | respiratory_support == '2,1') %>% nrow()
trans[2,1]<-filter(transmeta,d5 == 5) %>% 
  filter(respiratory_support != '1' & respiratory_support != '2,1') %>% 
   nrow()
trans[1,2]<-filter(transmeta,d5 != 5) %>% 
  filter(respiratory_support == '1' | respiratory_support == '2,1') %>% nrow()
trans[2,2]<-filter(transmeta,d5 != 5) %>% 
  filter(respiratory_support != '1' & respiratory_support != '2,1') %>% 
  nrow()
fisher.test(trans)

```

###severity cs234 cs6
```{r}
metadata = read.csv('F:/ZLF/CAP/data/relative_data/metadata-all-220617.csv',row.names = 1)
trans<-data.frame(matrix(nrow = 2,ncol = 2))
trans<-data.frame(matrix(nrow = 2,ncol = 2))
rownames(trans)<-c('severe','non-severe')
colnames(trans)<-c('cs234','cs6')

trans[1,1]<-filter(metadata,pam_10_cluster == 2 | pam_10_cluster ==3 | pam_10_cluster == 4) %>% 
  filter(severe_case == 'yes') %>% nrow()
trans[2,1]<-filter(metadata,pam_10_cluster == 2 | pam_10_cluster ==3 | pam_10_cluster == 4) %>% 
  filter(severe_case == 'no') %>% nrow()
trans[1,2]<-filter(metadata,pam_10_cluster == 6) %>% 
  filter(severe_case == 'yes') %>% nrow()
trans[2,2]<-filter(metadata,pam_10_cluster == 6) %>% 
  filter(severe_case == 'no') %>% nrow()
fisher.test(trans)
```

###severity cs234 cs15789
```{r}
metadata = read.csv('F:/ZLF/CAP/data/relative_data/metadata-all-220617.csv',row.names = 1)
trans<-data.frame(matrix(nrow = 2,ncol = 2))
trans<-data.frame(matrix(nrow = 2,ncol = 2))
rownames(trans)<-c('severe','non-severe')
colnames(trans)<-c('cs234','cs15789')

trans[1,1]<-filter(metadata,pam_10_cluster == 2 | pam_10_cluster ==3 | pam_10_cluster == 4) %>% 
  filter(severe_case == 'yes') %>% nrow()
trans[2,1]<-filter(metadata,pam_10_cluster == 2 | pam_10_cluster ==3 | pam_10_cluster == 4) %>% 
  filter(severe_case == 'no') %>% nrow()
trans[1,2]<-filter(metadata,pam_10_cluster == 1 | pam_10_cluster ==5 | pam_10_cluster == 7 | pam_10_cluster == 8 | pam_10_cluster == 9) %>% 
  filter(severe_case == 'yes') %>% nrow()
trans[2,2]<-filter(metadata,pam_10_cluster == 1 | pam_10_cluster ==5 | pam_10_cluster == 7 | pam_10_cluster == 8 | pam_10_cluster == 9) %>% 
  filter(severe_case == 'no') %>% nrow()
fisher.test(trans)
```



###CS day1 to day5
```{r}
  unique(transmeta$respiratory_support)
pva<-data.frame(cs = 1,trans_to_cs = 1,pvalue = NA)
transmeta = read.xlsx('F:/ZLF/CAP/data/relative_data/trans_meta1.xlsx',1,rowNames = T)
transmeta = read.csv('F:/ZLF/CAP/data/relative_data/trans_meta1.csv',row.names = 1)
transmeta<-filter(transmeta,severity == 'yes')
#transmeta<-filter(transmeta,respiratory_support == '1' | respiratory_support == '2,1')
transmeta<-transmeta[!is.na(transmeta[,'d1']),]
transmeta<-transmeta[!is.na(transmeta[,'d5']),]
transmeta<-filter(transmeta,severity == 'yes'|severity == 'no')
j = 8
i = 5

#View(transmeta)

#transmeta$severity

#transmeta$severity
for (j in 1:10) {
  for (i in 1:10) {
    #assign(paste0('trans',j,i))
    trans<-data.frame(matrix(nrow = 2,ncol = 2))
    trans<-data.frame(matrix(nrow = 2,ncol = 2))
    rownames(trans)<-c('cs','csothers')
    colnames(trans)<-c('transtocs','transtoothers')
    
    trans[1,1]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == j & d5 ==i))
    trans[2,1]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 != j & d1 != i & d5 == i))
    trans[1,2]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == j & d5 != i &d5 != j))
    mid = 0
    #for (k in setdiff(setdiff(1:10,j),i)) {for(h in setdiff(setdiff(1:10,i),k)){mid = mid+nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == k & d5 ==h))}}
    for (k in setdiff(1:10,j)) {for(h in setdiff(setdiff(1:10,i),k)){mid = mid+nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == k & d5 ==h))}}
    
    trans[2,2] = mid
    #trans[2,2]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 != j & d5 != i))
    pva[10*(as.numeric(j)-1)+as.numeric(i),'trans_to_cs'] = as.numeric(i)
    pva[10*(as.numeric(j)-1)+as.numeric(i),'cs'] = as.numeric(j)
    pva[10*(as.numeric(j)-1)+as.numeric(i),'pvalue'] = fisher.test(trans)$p
    #print(paste('cluter',j,'trans to',i,'pvalue = ',fisher.test(trans)$p))
  }                
}

pva$qvalue<-p.adjust(pva$pvalue,method = 'bh')
pva_sig <- filter(pva,pvalue <= 0.05) %>% filter(cs != trans_to_cs)

j = 8
i = 5
trans<-data.frame(matrix(nrow = 2,ncol = 2))
trans<-data.frame(matrix(nrow = 2,ncol = 2))
rownames(trans)<-c('cs','csothers')
colnames(trans)<-c('transtocs','transtoothers')
trans[1,1]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == j & d5 ==i))
trans[2,1]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 != j & d1 != i & d5 == i))
trans[1,2]<-nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == j & d5 != i &d5 != j))
mid = 0
for (k in setdiff(1:10,j)) {for(h in setdiff(setdiff(1:10,i),k)){mid = mid+nrow(filter(transmeta[!is.na(transmeta[,'d1']),],d1 == k & d5 ==h))}}
trans[2,2] = mid


rownames(trans)[1] <- c('cs8')
colnames(trans)[1] <- 'transtocs5'

trans

fisher.test(trans)


```



